### 2012-05-10 Rctl/inst/doc/manual.Rnw
[Rctl] Manual updates

After you have stored the data as plain text open the files in a text editor (NOT WORD, for 
windows Notepad++ is a good option) and check to see if your data is stored correctly, if 
not reformat your data in both the genotype.csv and the phenotype.csv files to match the 
following:

<<echo=TRUE,eval=FALSE>>=
setwd("<PATH_TO_DATA>")
genotypes <- read.csv("genotypes.csv",row.names=1,col.names=TRUE)
traits <- read.csv("traits.csv",row.names=1,col.names=TRUE)
genotypes[1:5,1:10] #Show the first 5 individuals, 10 markers
traits[1:5,1:10] #Show the first 5 individuals, 10 traits
@

@@

@@

Some more advanced topics

>>>>

After you have stored the data as plain text open the files in a text editor (NOT WORD, for 
windows Notepad++ is a good option) and check to see if your data is stored correctly, if 
not reformat your data in both the genotype.csv and the phenotype.csv files to match the 
format below. {\color{red}\emph{\bf{NOTE: Individuals must match between the genotype and 
the phenotype file...}}} The phenotypes.csv file containing the genotype matrix is stored 
individuals x genetic marker:

\begin{tabular}{lrrrrr}
      & Marker1 & Marker2 & Marker3 & ... & MarkerN \\
Ind1  & 1 & 1 & 2 & ... & 2 \\
Ind2  & 1 & 1 & 2 & ... & 2 \\
Ind3  & 1 & 1 & 1 & ... & 2 \\
...   & ... & ... & ... & ... & ... \\
IndN  & 1 & 1 & 1 & ... & 2 \\
\end{tabular}

The phenotypes.csv file containing individuals x trait measurements:

\begin{tabular}{lrrrrr}
      & Trait1 & Trait2 & Trait3 & ... & TraitN \\
Ind1  & 5.8    & 11.0   & 9.0    & ... & 1.6 \\
Ind2  & 6.3    & 12.2   & NA     & ... & 1.3 \\
Ind3  & 5.1    & 11.1   & 12.3   & ... & 2.0 \\
...   & ...    & ...    & ...    & ... & ... \\
IndN  & 9.8    & 15.8   & 23.0   & ... & 3.4 \\
\end{tabular}

After verifying your files are formatted correctly (this is the source of most errors).
Start R and load in the data, check that the output looks like below:
<<echo=TRUE,eval=FALSE>>=
setwd("<PATH_TO_DATA>")
genotypes <- read.csv("genotypes.csv",row.names=1,col.names=TRUE)
traits <- read.csv("traits.csv",row.names=1,col.names=TRUE)
genotypes[1:5,1:10] #Show the first 5 individuals, 10 markers
traits[1:5,1:10] #Show the first 5 individuals, 10 traits
@

@@

\emph{Adding genetic map information}
This step is optional, but plots and generated networks look much nicer when chromosome locations
of genetic markers are supplied. The 'mapinfo' object is matrix with 3 columns: "Chr" - the 
chromosome number, "cM" - the location of the marker in centiMorgans and the 3rd column "Mbp" - The 
location of the marker in Mega basepairs. The structure of the $mapinfo.csv$ file:

\begin{tabular}{lrrr}
      & Chr    & cM     & Mbp \\
Ind1  & 1    & 1.0   & 0.01   \\
Ind2  & 1    & 1.2   & 0.34   \\
Ind3  & 1    & 3.1   & 1.3    \\
...   & ...  & ...   & ...    \\
IndN  & 24   & 120.8 & 20.3   \\
\end{tabular}

We load the genetic map data into R by using the following command:

<<echo=TRUE,eval=FALSE>>=
mapinfo <- read.csv("mapinfo.csv",row.names=1,col.names=TRUE)
traits[1:5,1:3] #Show the first 5 marker records
@
After this we can
\bigskip

@@

So now that we have our ctls calculated we can visualize the results in multiple ways.
The easiest way is first to plot the two overview CTL heatmaps. The first heatmap shows 
the summarized LOD scores for the Trait X Marker CTLs. This plot normally looks very 
similar to a QTL heatmap:
<<echo=TRUE,eval=FALSE>>=
image(ctls,against="markers")
@
The second heatmap shows the summarized LOD scores for the Trait * Trait CTLs, this 
shows the summarized amount of evidence we find for each Trait * Trait interaction:
<<echo=TRUE,eval=FALSE>>=
image(ctls,against="phenotypes")
@
After the overview plots it is time to go more in depth, and look at some individual CTL 
profiles the CTLobject 'ctls' contains multiple CTLscans. To print a summary of the first 
scan use:
<<echo=TRUE,eval=FALSE>>=
ctls[[1]]
@
To plot other summaries just replace the 1 by the trait number. We can also use the plot 
function on individual CTLscans and shows a colorful and detailed view of the 
(selected) Trait x Marker X Traits interactions detected in the dataset.
<<echo=TRUE,eval=FALSE>>=
plot(ctls[[1]])
@
